---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# distspec: probability distributions with certain or uncertain parameters

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![R-CMD-check](https://github.com/epiforecasts/distspec/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/epiforecasts/distspec/actions/workflows/R-CMD-check.yaml) [![Codecov test coverage](https://codecov.io/gh/epiforecasts/distspec/graph/badge.svg)](https://app.codecov.io/gh/epiforecasts/distspec) [![CRAN status](https://www.r-pkg.org/badges/version/distspec)](https://CRAN.R-project.org/package=distspec)

[![MIT license](https://img.shields.io/badge/License-MIT-blue.svg)](https://github.com/epiforecasts/distspec/blob/main/LICENSE.md) [![GitHub contributors](https://img.shields.io/github/contributors/epiforecasts/distspec)](https://github.com/epiforecasts/distspec/graphs/contributors) [![universe](https://epiforecasts.r-universe.dev/badges/distspec)](https://epiforecasts.r-universe.dev/distspec)
<!-- badges: end -->

distspec represents a probability distribution as a single object, a
`<dist_spec>`, whose parameters can be either fixed or uncertain. It grew out of
[EpiNow2](https://epiforecasts.io/EpiNow2/) and is aimed at the delay
distributions common in infectious disease modelling, such as generation times,
incubation periods and reporting delays, while staying independent of any
particular model.

With distspec you can:

- define a distribution with a named constructor (`Gamma()`, `LogNormal()`,
  `Normal()`, `Exponential()`, `Weibull()`, `Beta()`, `Fixed()` or
  `NonParametric()`),
  by its natural parameters or by its mean and standard deviation;
- give any parameter an uncertain prior (for example
  `Gamma(shape = Normal(2, 0.5), rate = 1)`), or leave a nonparametric mass
  function uncertain via a `Dirichlet()` prior;
- discretise a continuous distribution to a probability mass function
  (`discretise()`), convolve distributions (`+`, `collapse()`), draw samples
  (`sample_dist()`), and query means, standard deviations, bounds and PMFs.

## Installation

Install the development version from GitHub:

```r
# install.packages("remotes")
remotes::install_github("epiforecasts/distspec")
```

## Quick start

```{r example}
library(distspec)

# A gamma delay with mean 4 and standard deviation 2, truncated at 20
delay <- Gamma(mean = 4, sd = 2, max = 20)
delay

# Discretise it to a probability mass function
get_pmf(discretise(delay))

# A parameter can itself be a distribution, expressing uncertainty
Gamma(shape = Normal(2, 0.5), rate = 1)
```

`plot()` shows the probability mass and cumulative distribution functions:

```{r plot, fig.alt = "PMF and CDF of a gamma delay with mean 4 and standard deviation 2."}
library(ggplot2)
plot(delay)
```

See `vignette("distspec")` to get started, and the
[reference index](https://epiforecasts.io/distspec/reference/) for the full list
of functions.

## Related work

distspec discretises using
[primarycensored](https://primarycensored.epinowcast.org/), which implements the
double-censoring calculation. In Julia, the same censoring maths lives in
[CensoredDistributions.jl](https://github.com/EpiAware/CensoredDistributions.jl),
and [ComposedDistributions.jl](https://composeddistributions.epiaware.org/)
covers similar ground to the `<dist_spec>` combination interface, with
`compose()` / `sequential()` in place of `+` and `observed_distribution` in
place of `collapse()`. Both distspec and ComposedDistributions.jl carry
parameter uncertainty in the object. The Julia packages build on
[Distributions.jl](https://juliastats.org/Distributions.jl/) and represent
richer event trees, while distspec keeps a flatter, R-native representation.

Contributors
---





<!-- ALL-CONTRIBUTORS-LIST:START - Do not remove or modify this section -->
<!-- prettier-ignore-start -->
<!-- markdownlint-disable -->

All contributions to this project are gratefully acknowledged using the [`allcontributors` package](https://github.com/ropensci/allcontributors) following the [all-contributors](https://allcontributors.org) specification. Contributions of any kind are welcome!

### Code


<a href="https://github.com/epiforecasts/distspec/commits?author=sbfnk">sbfnk</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=dependabot[bot]">dependabot[bot]</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=github-merge-queue[bot]">github-merge-queue[bot]</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=seabbs">seabbs</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=github-actions[bot]">github-actions[bot]</a>



### Issues


<a href="https://github.com/epiforecasts/distspec/issues?q=is%3Aissue+author%3Ajamesmbaazam">jamesmbaazam</a>


<!-- markdownlint-enable -->
<!-- prettier-ignore-end -->
<!-- ALL-CONTRIBUTORS-LIST:END -->
