---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# distspec: probability distributions with certain or uncertain parameters

<!-- badges: start -->
[![R-CMD-check](https://github.com/epiforecasts/distspec/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/epiforecasts/distspec/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/epiforecasts/distspec/graph/badge.svg)](https://app.codecov.io/gh/epiforecasts/distspec)
<!-- badges: end -->

distspec represents a probability distribution as a single object, a
`<dist_spec>`, whose parameters can be either fixed or uncertain. It grew out of
[EpiNow2](https://epiforecasts.io/EpiNow2/) and targets the delay distributions
that recur in infectious disease modelling — generation times, incubation
periods, reporting delays — while remaining independent of any particular model.

With distspec you can:

- define a distribution with a named constructor (`Gamma()`, `LogNormal()`,
  `Normal()`, `Exp()`, `Weibull()`, `Beta()`, `Fixed()` or `NonParametric()`),
  by its natural parameters or by its mean and standard deviation;
- give any parameter an uncertain prior (for example
  `Gamma(shape = Normal(2, 0.5), rate = 1)`), or leave a nonparametric mass
  function to be estimated during model fitting via a `Dirichlet()` prior;
- discretise a continuous distribution to a probability mass function
  (`discretise()`), convolve distributions (`+`, `collapse()`), draw samples
  (`sample_dist()`), and query means, standard deviations, bounds and PMFs.

## Installation

Install the development version from GitHub:

```r
# install.packages("remotes")
remotes::install_github("epiforecasts/distspec")
```

## Quick start

```{r example}
library(distspec)

# A gamma delay with mean 4 and standard deviation 2, truncated at 20
delay <- Gamma(mean = 4, sd = 2, max = 20)
delay

# Discretise it to a probability mass function
get_pmf(discretise(delay))

# A parameter can itself be a distribution, expressing uncertainty
Gamma(shape = Normal(2, 0.5), rate = 1)
```

See `vignette("distspec")` to get started, and the
[reference index](https://epiforecasts.io/distspec/reference/) for the full list
of functions.

Contributors
---





<!-- ALL-CONTRIBUTORS-LIST:START - Do not remove or modify this section -->
<!-- prettier-ignore-start -->
<!-- markdownlint-disable -->

All contributions to this project are gratefully acknowledged using the [`allcontributors` package](https://github.com/ropensci/allcontributors) following the [all-contributors](https://allcontributors.org) specification. Contributions of any kind are welcome!

### Code


<a href="https://github.com/epiforecasts/distspec/commits?author=sbfnk">sbfnk</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=dependabot[bot]">dependabot[bot]</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=github-merge-queue[bot]">github-merge-queue[bot]</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=seabbs">seabbs</a>,
<a href="https://github.com/epiforecasts/distspec/commits?author=github-actions[bot]">github-actions[bot]</a>



### Issues


<a href="https://github.com/epiforecasts/distspec/issues?q=is%3Aissue+author%3Ajamesmbaazam">jamesmbaazam</a>


<!-- markdownlint-enable -->
<!-- prettier-ignore-end -->
<!-- ALL-CONTRIBUTORS-LIST:END -->
