---
title: "Adding a distribution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding a distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This guide is for contributors adding a new distribution to distspec. It sets
out what a distribution consists of, which S3 methods it must define, and which
it defines only when it supports a given capability. A completeness test
(`tests/testthat/test-completeness.R`) enforces the core contract, so a
half-added distribution fails CI.

## The anatomy of a `<dist_spec>`

Every distribution in distspec is a `<dist_spec>` object. The user-facing
constructors ([`Gamma()`], [`LogNormal()`] and the rest) all hand their
arguments to `new_dist_spec()`, which builds the object. A single parametric
`<dist_spec>` is a list with two fields:

- `parameters`: a named list of the distribution's parameters, each either a
  fixed number or another `<dist_spec>` (an uncertain parameter, i.e. a prior);
- `distribution`: the type as a string, for example `"gamma"`.

Its class is `c("<type>", "dist_spec")`. Per-type S3 methods dispatch on the
`"<type>"` head; shared behaviour that is the same for every distribution
(printing, addition, discretisation) dispatches on the `"dist_spec"` tail. An
uncertain distribution additionally carries an `"uncertain"` class at
the front so the shared uncertainty handlers pick it up.

`new_dist_spec()` does the shared work once for every type: it checks that the
right number of parameters was given, validates them against their lower bounds,
converts any non-natural parameters (such as a mean and standard deviation) to
the natural parameters, propagates uncertainty, and applies any bounds. A new
distribution therefore does not repeat any of that. It supplies only the
per-type pieces below.

## One file per distribution

Each distribution lives in a single file `R/<type>.R` that holds its constructor
and all of its S3 methods. For example `R/gamma.R` contains `Gamma()` together
with `natural_params.gamma()`, `lower_bounds.gamma()` and so on. Add your new
distribution the same way, in its own `R/<type>.R`.

## The constructor

The constructor collects its arguments and calls `new_dist_spec()` with the
parameter list and the type string. It accepts both the natural parameters and
any convenient alternatives (such as `mean`/`sd`), and passes `...` through for
the bounds:

```{r}
MyDist <- function(rate, mean, ...) {
  params <- as.list(environment())
  new_dist_spec(params, "mydist", ...)
}
```

`new_dist_spec()` reads `natural_params()` to work out how many parameters are
required, so exactly one valid parameterisation must be supplied.

## The core contract

Every distribution must define these three S3 methods. The completeness test
checks that each of them exists for every registered type.

- `natural_params.<type>(x)`: returns a character vector naming the natural
  (canonical) parameters, for example `c("shape", "rate")`. This drives
  parameter validation and conversion, so it is what makes a type a
  distribution.
- `lower_bounds.<type>(x)`: returns a named numeric vector of lower bounds, one
  per parameter (natural and alternative), used to reject unsupported values.
- `mean.<type>(x, ...)`: returns the mean of the distribution from its (fixed)
  natural parameters.

```{r}
natural_params.mydist <- function(x) "rate"

lower_bounds.mydist <- function(x) c(rate = 0, mean = 0)

mean.mydist <- function(x, ...) 1 / x$parameters$rate
```

## The conditional contract

The remaining methods are defined only when the distribution supports the
matching capability. Define each one when its condition holds.

- `to_natural.<type>(x, ux)`: define this if the distribution can be given by
  parameters other than its natural ones (for example a mean and standard
  deviation). It maps the supplied parameter means `ux` to a named list of
  natural parameters. A distribution that only ever takes its natural parameters
  (such as the fixed point mass) omits it.
- `sd.<type>(x, ...)`: define this if the distribution has a scalar standard
  deviation. Distributions whose "spread" is not a single number (such as the
  Dirichlet prior) omit it.
- `sample_dist.<type>(x, n, ...)`: define this if the distribution can be sampled
  directly, usually by calling the base-R generator for its family (`rgamma()`,
  `rlnorm()` and so on). It draws `n` samples from the fixed distribution.
- `dist_cdf.<type>(x)`: define this if the distribution is continuous and should
  be discretisable through its cumulative distribution function. It returns the
  base-R `p*` function (for example `pgamma`). A type without a `dist_cdf()`
  method cannot be discretised through a CDF and errors informatively.

Two capabilities are provided by overriding a shared method rather than a
per-type generic, so they are needed only in special cases:

- `discrete_pmf.<type>(x, max_value, ...)`: override this only if the
  distribution discretises to a PMF by some route other than a CDF (the fixed
  point mass does this).
- `max.<type>(x, ...)`: override this only if the support size is not read from
  parameters in the usual way (the nonparametric distribution does this).

## A worked template

A file skeleton for a fictional single-parameter distribution `MyDist`, which
supports every capability, would look like this. Place it in `R/mydist.R`:

```{r}
# My distribution
#
# Per-type methods for the fictional MyDist distribution.

#' My distribution
#'
#' @param rate Rate of the distribution.
#' @param mean Mean of the distribution, as an alternative to `rate`.
#' @param ... Limits of the distribution, passed to [bound_dist()].
#' @return A `<dist_spec>`.
#' @export
MyDist <- function(rate, mean, ...) {
  params <- as.list(environment())
  new_dist_spec(params, "mydist", ...)
}

# --- core contract ---

#' @exportS3Method
natural_params.mydist <- function(x) "rate"

#' @exportS3Method
lower_bounds.mydist <- function(x) c(rate = 0, mean = 0)

#' @method mean mydist
#' @export
mean.mydist <- function(x, ...) 1 / x$parameters$rate

# --- conditional contract ---

# given by mean instead of rate: define to_natural()
#' @exportS3Method
to_natural.mydist <- function(x, ux) {
  list(rate = if ("mean" %in% names(ux)) 1 / ux$mean else ux$rate)
}

# has a scalar sd: define sd()
#' @method sd mydist
#' @export
sd.mydist <- function(x, ...) 1 / x$parameters$rate

# can be sampled: define sample_dist()
#' @importFrom stats rexp
#' @exportS3Method
sample_dist.mydist <- function(x, n, ...) {
  rexp(n, rate = x$parameters$rate)
}

# continuous and discretisable: define dist_cdf()
#' @exportS3Method
dist_cdf.mydist <- function(x) pexp
```

Run `devtools::document()` after adding the roxygen tags so the S3 methods are
registered in `NAMESPACE`.

## The completeness test

`tests/testthat/test-completeness.R` enumerates every type that registers a
`natural_params()` method and asserts that each also defines the rest of the
core contract. If you register `natural_params.<type>()` without the other core
methods, the test fails, which flags a half-added distribution before it is
merged. Extend the test if your distribution genuinely cannot satisfy part of
the core contract, and say why in the file.
