---
title: "Getting started with distspec"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with distspec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(1)
```

```{r setup}
library(distspec)
```

In distspec, a probability distribution is a single object: a `<dist_spec>`. It
can hold fixed parameters or uncertain ones, and the same functions work on
either.

## Defining a distribution

Each distribution has its own constructor. Give it the natural parameters, or a
mean and standard deviation that distspec converts for you:

```{r define}
Gamma(shape = 2, rate = 0.5)
Gamma(mean = 4, sd = 2)
LogNormal(meanlog = 1, sdlog = 0.5)
```

A finite maximum (and, for parametric distributions, a `cdf_cutoff`) truncates
the support:

```{r bound}
Gamma(mean = 4, sd = 2, max = 20)
```

## Uncertain parameters

Any parameter can be a number or, to express uncertainty about its value,
another `<dist_spec>`. Uncertain parameters must be given as the natural
parameters:

```{r uncertain}
uncertain <- Gamma(shape = Normal(2, 0.5), rate = Normal(0.5, 0.1))
uncertain

# the mean of an uncertain distribution is unknown unless we ignore uncertainty
mean(uncertain)
mean(uncertain, ignore_uncertainty = TRUE)
```

`fix_parameters()` resolves an uncertain distribution to a fixed one, taking
either the mean of each prior or a sample from it:

```{r fix}
fix_parameters(uncertain, strategy = "mean")
```

## Discretising

`discretise()` turns a continuous distribution into a nonparametric probability
mass function over `0, 1, 2, ...`:

```{r discretise}
pmf <- discretise(Gamma(mean = 4, sd = 2, max = 20))
get_pmf(pmf)
```

## Combining distributions

Adding distributions convolves them; `collapse()` reduces a composite to a
single discretised PMF:

```{r combine}
combined <- Gamma(mean = 4, sd = 2, max = 20) +
  LogNormal(meanlog = 1, sdlog = 0.5, max = 20)
mean(collapse(discretise(combined)))
```

## Sampling

`sample_dist()` draws random samples from a distribution with fixed parameters:

```{r sample}
sample_dist(Gamma(mean = 4, sd = 2, max = 20), n = 5)
```

## Estimated nonparametric distributions

Instead of a fixed PMF, a nonparametric distribution can carry a `Dirichlet()`
prior over its bins, to be estimated during model fitting:

```{r estimated}
est <- NonParametric(pmf = Dirichlet(c(0, 2, 4, 3)))
est
```

It then has no concrete PMF until you resolve it with `fix_parameters()`.
`has_uncertainty()` reports whether a distribution carries a prior:

```{r has-uncertainty}
has_uncertainty(est)
has_uncertainty(Gamma(shape = 2, rate = 0.5))
```
